---
title: "Time series Analysis"
output:
  output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(ggplot2)
library(tidyverse) 
library(tseries)

```

## Preparation of the data
We begin by loading our data and convert it to a time series using the ``as.ts()`` function. Then we plot our data.
```{r,echo=FALSE}
filepath="~/Desktop/MSc-Imperial/Term 2/Data visualisation/Assessments/Assessment_2/time-series.csv"
```

```{r,out.width='75%',fig.align = 'center'}
ts <- read_csv(filepath,show_col_types = FALSE)
tsData<-as.ts(ts) ##convert to time series data
plot.ts(tsData)
grid(nx = NA,
     ny = NULL,
     lty = 2, col = "gray", lwd = 2)
```
It's very hard to distinguish if there is a trend or seasonality by looking at the time series graph.We proceed by looking at the ACF plot.
```{r,out.width='75%',fig.align = 'center'}
acf(tsData) # autocorrelation
```
We can observe that there is a trend and seasonal component in our time series.The slow decrease in the ACF as the lags increase is due to the trend component, while the periodic sinoid shape is due the seasonality.

## Parameters for the linear trend component
We fit a linear model by regressing our series ``ts`` against time and plot the regression line. The upward trend becomes clearer in the new plot. 

```{r,out.width='65%',fig.align = 'center'}
##Linear regression
trModel<- lm(tsData ~ time(tsData))
##Plot time series with trending line
plot.ts(tsData)
grid(nx = NA,
     ny = NULL,
     lty = 2, col = "gray", lwd = 2)
abline(reg=trModel)
```

We remove the trend by computing the first difference.The fitted values will be the estimates ``t`` for the trend component. In the new graph below we can observe a shift; the mean of our values are around the x-axis so there's no need to compute another difference again. 

```{r,out.width='65%',fig.align = 'center'}
##Removing the trend
trend <- trModel$fitted.values
d.data <- tsData - trend
##Plot the data
plot(d.data)
grid(nx = NA,
     ny = NULL,
     lty = 2, col = "gray", lwd = 2)
```
The parameters for the linear trend component are:
```{r,echo=FALSE}
##Removing the trend
knitr::kable(trModel$coefficients, "simple")
```


## Seasonal componen

To interpret the seasonality and determine the frequency, we plot the ACF and PACF after removing the linear trend.
```{r,echo=FALSE,out.width='65%',fig.align = 'center'}
acf(d.data) # autocorrelation
pacf(d.data)
```
Looking at the ACF plot, we observe that there is a periodic trend with positive and negative spikes following each other. The highest peak occurs at lag 28 and there is another less significant peak at lag 14.The PACF is also kind of periodic and has some significant spikes from lag 3 to 7 and some small spikes coming after lag 15. Then, it cuts off after lag 24.

We are going to perform spectral analysis  to get an idea of the frequency and the periodic behavior.To learn where the dominant peak is located, we assign a data frame to the spectrum output.We find the frequency ``f`` at which the corresponding spectral density is largest and then we compute the period ``T`` corresponding to ``f`` by ``T=1/f``. There is a large spike at frequency ``0.00357``, which comes out to a cycle with a period of $1/0.00357 \sim 28$ .

```{r,out.width='65%',fig.align = 'center'}
spec<-spectrum(d.data)
spec.df <- data.frame(spec=spec$spec, freq=spec$freq)
frequency<-spec$freq[spec$spec==max(spec$spec)]
period<-round(1/frequency)
```
```{r,echo=FALSE}
df <- data.frame (frequency,period)
knitr::kable(df, "pipe")
```
We determine that there is a period of 28 lags which we will use to remove the seasonal component from the original data. In order to do this we compute the difference $y_t-S_t$, where $S_t=S_{28}$. 
```{r}
diff.ts=diff(d.data, lag=T)  #lag difference 
```

```{r,echo=FALSE,out.width='55%',fig.align = 'center'}
acf(diff.ts)
pacf(diff.ts)
```
In the ACF, there are two significant spikes at lag 1 and 2 and a less significant at lag 3.Then we observe a decreasing trend in values, although there is a cluster of less significant negative and positive spikes around 14 and 28. This indicates a pattern around multiples of 14.
The auto correlations in PACF are smaller in scale.There is a clear spike at lag 1, 2 and 3 and a cluster of negative spikes around lag 14. 
The spikes in the PACF around lag 14 and 28 combined with spikes around the same lags in the ACF indicates that a four-weekly seasonal component may exist. Moreover, the spikes in low lags in the PACF in combination with the decreasing trend of the ACF plot suggests a possible non seasonal AR.

## Dickey-Fuller test
After removing the trending and seasonal components, we're going to test if the time series is stationary.We're using a lag order of k=1,for the simple Dickey-Fuller test. 

```{r}
adf.test(diff.ts, k=1)
```
We see that the p-value is (smaller than) 0.01, which suggests that we can reject the null hypothesis and conclude that the time series is stationary.However, we can't claim that this is evidence in support of a particular alternative hypothesis (e.g. stationarity)

## Suitability of test
The Dickey–Fuller test that we used above is used to test the null hypothesis, which is that a unit root is present in an autoregressive time series model.However, it can not detect other forms of nonstationarity as the alternative hypothesis is different depending on which version of the test is used.  
Our time series is clearly seasonal and more complicated than is captured by a simple AR(1) model. A more appropriate test would be the augmented Dickey Fuller test which was developed for general ARMA(p, q) models
with unknown orders. This seems to be closer to our case and thus more appropriate. 


